---
title: "Netflix Case Study"
author: "marty"
date: "2025-02-18"
output:
  pdf_document: default
  html_document: default
---

## Introduction

The following analysis is a case study about Netflix. This analysis aims to answer the question: **What differentiates Netflix's content from other streaming services?**

Ultimately, answering this question can help move the needle with a core problem in the streaming industry: capturing subscribers

Throughout the study, I will be using data on Netflix's content from 2016-2021 alongside a database from TMDB (The Movie Database). The TMDB database contains 1.2 million observations of historical data about movies and TV shows (e.g.genres, keywords, popularity). I'll be manipulating that data to find any trends in Netflix's content over the years, and thus identify what makes them different from their competitors. 

These are the datasets I'll be using:

* [Netflix Movies & TV Shows](https://www.kaggle.com/datasets/shivamb/netflix-shows): created by Shivam Bansal, under the [CC0: Public Domain](https://creativecommons.org/publicdomain/zero/1.0/) license
* [Full TMDB Movies Database 2024](https://www.kaggle.com/datasets/asaniczka/tmdb-movies-dataset-2023-930k-movies/data): created by user name asaniczka on Kaggle, under the [ODC Attribution](https://opendatacommons.org/licenses/by/1-0/index.html) license
* [Full TMDB TV Shows Database 2024](https://www.kaggle.com/datasets/asaniczka/full-tmdb-tv-shows-dataset-2023-150k-shows): created by user name asaniczka on Kaggle, under the [ODC Attribution](https://opendatacommons.org/licenses/by/1-0/index.html) license

## Setting up the Environment 

To start, I'll be installing the tidyverse library of functions for its data analysis capabilities.

```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)
```

## Importing the Datasets

Then, I'll be importing the datasets under the following aliases on RStudio:

1. *netflix*: a data frame containing all movies and TV shows on the Netflix streaming service from 2016-2021
2. *movies*: a data frame containing 1.1 million observations on movies, including details such as genre, keywords, audience score, budget, etc.
3. *shows*: a data frame containing 168,000 observations on TV shows, including details such as genre, descriptions, release date, etc.

```{r}
# importing the netflix data set as "netflix"
netflix <- read.csv("C:\\Users\\marti\\Desktop\\NETFLIX CASE STUDY B\\Data Sources\\netflix_titles.csv")

# importing the movies data set in a "movies" table
movies <- read.csv("C:\\Users\\marti\\Desktop\\NETFLIX CASE STUDY B\\Data Sources\\movie_dataset.csv")

# importing the shows data as in a "shows" table
shows <- read.csv("C:\\Users\\marti\\Desktop\\NETFLIX CASE STUDY B\\Data Sources\\tv_datset.csv")
```

## Preparing the Datasets

Not every variable in the **movies** and **shows** tables was going to be useful in the analysis, so I removed the unnecessary ones below:

```{r}
# removing unnecessary or redundant variables from the movies table
movies <- movies %>% select(-spoken_languages,-production_countries,-tagline,-poster_path,-overview,-original_title,-original_language,-imdb_id,-homepage,-backdrop_path,-adult,-runtime,-status,-id)

# removing unnecessary or redundant variables from the shows table
shows <- shows %>% select(-episode_run_time,-production_countries,-spoken_languages,-origin_country,-networks,-overview,-languages,-tagline,-status,-type,-poster_path,-original_name,-in_production,-homepage,-backdrop_path,-adult,-original_language,-number_of_seasons,-id)
```

## Cleaning the Datasets

Next, I needed to join the **netflix** table with the other two so that I could analyze trends throughout Netflix's content over the years. All three tables had unique ID values for its movies/shows, so I decided to *join the tables based on  their titles (title) and release data (date_added).* Titles were already set for all the tables, so I just had to standardize the dates.

### Fixing the date_added column of the netflix table
```{r}
# Change the format of date added from a chr to a date time object
netflix$date_added <- mdy(netflix$date_added)
```

### Fixing the date column of the movies table
```{r}

#change the format of the release date column in movies from chr to a date time object
movies$release_date <- ymd(movies$release_date)

#show only the year for the release date column (in order to match the values in the netflix release year column)
movies$release_date <- year(movies$release_date)

#rename the release date column in movies to enable joining with the netflix table
movies <- movies %>% rename(release_year=release_date)
```

With dates standardized across the netflix and movies table, I could now join them together:

### Creating a "netmovies" table that joins the netflix and movies tables for additional movies details
```{r}
# Join the movies and netflix table according to movie title and release date
netmovies <- merge(x=netflix,y=movies,by=c("title","release_year"),all.x=TRUE)

# Remove duplicate entries by only keeping the unique movies (filter by id and content type of the netflix table)
netmovies <- netmovies %>% distinct(show_id,.keep_all=TRUE) %>% filter(type=="Movie")

# Adding a year released column
netmovies <- netmovies %>% mutate(year_added = year(date_added))
```

### Note: There are also null values for a number of movies (1545 fields) on additional variables such as **vote_average, vote_count, revenue, budget, & popularity**, but I chose to keep these records in the analysis because of the **genre, year_added, and rating** variables

Next, I had to join the netflix and shows table:

### Preparing the shows table
```{r}
# 
#change the format of the release date column in shows from chr to a date time object
shows$first_air_date <- ymd(shows$first_air_date)

#show only the year for the release date column (in order to match the values in the netflix release year column)
shows$first_air_date <- year(shows$first_air_date)

#rename the title and date columns in the shows table to enable joining with the netflix table
shows <- shows %>% rename(release_year=first_air_date,title=name)
```

### Creating a "netshows" table that joins the netflix and shows tables for additional show details
```{r}
# Join the shows and netflix table according to show name and release date
netshows <- merge(x=netflix,y=shows,by=c("title","release_year"),all.x=TRUE)

# Remove duplicate entries by only keeping the unique show ids (and TV shows of the netflix table)
netshows <- netshows %>% distinct(show_id,.keep_all=TRUE) %>% filter(type=="TV Show")

# Adding a year released column
netshows <- netshows %>% mutate(year_added = year(date_added))
```

With that being done, I can now conduct my analysis on these two primary tables: *netmovies* and *netshows*

## Analysis

### Distribution of Genres over Time

Plotting the genres of Netflix's content over time can be a general measure of users' entertainment preferences. It can show which genres perform well, are growing, or remain the same. Some quick insights on the plot below:

* *Crime/Comedy/Drama seems to be Netflix's "bread and butter" genres* (marked in yellow and green) for movies, since it comprises more than 60% of their releases every year
* 2017-2019 saw the most growth for content released by Netflix. Within that time period, Netflix added *Drama (+216 movies), Comedy (+195), and Romance (+127) movie genres the most.*

```{r}
# Plot the genres of all movies throughout the years for Netflix
netmovies %>% separate_rows(genres) %>% filter(genres != "" & year_added > 2015 & year_added < 2022) %>% group_by(year_added,genres) %>% summarize(count=n()) %>% arrange(desc(count)) %>% ggplot(aes(x=year_added,y=count,fill=genres)) + geom_area(stat="identity") + xlim(2016,2021) + labs(title="Netflix Movie Genre Distribution from 2016-2021")

# Observing which genres saw the most growth for movies
growthmovies <- netmovies %>% separate_rows(genres) %>% filter(genres != "" & year_added > 2016 & year_added < 2020) %>% group_by(genres, year_added) %>% summarize(count=n()) %>% print(n=70)

```
Looking at the plot for TV Shows, 

**Drama was a top priority*, accounting for the most releases out of all genres every year
* 2017-2019 saw the most growth for content released by Netflix. Within that time period, *Drama (+80 shows), Sci-Fi (+27), and Crime (+25) saw the biggest growth amongst all TV shows*

```{r}

# Plot the genres of all shows throughout the years

netshows %>% separate_rows(genres) %>% filter(genres != "" & year_added > 2015 & year_added < 2022) %>% group_by(year_added,genres) %>% summarize(count=n()) %>% ggplot(aes(x=year_added,y=count,fill=genres)) + geom_area(stat="identity") + xlim(2016,2021) + labs(title="Netflix TV Shows Genre Distribution from 2015-2021")

# Observing which genres saw the most growth for tv shows 
growthtv <- netshows %>% separate_rows(genres) %>% filter(genres != "" & year_added > 2017 & year_added < 2020) %>% group_by(genres, year_added) %>% summarize(count=n()) %>% print(n=40)
```

### Ratings for Content

Generally, ratings are an indicator of the *minimum** appropriate age demographic for a form of entertainment. Analyzing the ratings of Netflix's content over the years may reveal things about its target market (age-wise). 

Using Netflix's [maturity ratings](https://help.netflix.com/en/node/2064) as a guide, *a large percentage of their content (43% of TV shows and 46% of movies) are recommended for adults.*

```{r}
# A general overview of the most common ratings on Netflix movies 
movieratings <- netmovies %>% group_by(rating) %>% summarize(movies=n()) %>% arrange(desc(movies)) %>% filter(rating != "" & movies > 2) 
movieratings <- movieratings %>% mutate(percentage_of_total=round((movies/sum(movies))*100,digits=2))
```

```{r}
# A general overview of the most common ratings on Netflix TV shows
showratings <- netshows %>% filter(rating != "") %>% group_by(rating) %>% summarize(shows=n()) %>% arrange(desc(shows)) 
showratings <- showratings %>% mutate(percentage_of_total=round((shows/sum(shows))*100,digits=2))
```

Identifying the characteristics of an adult-rated movie:

```{r}
# What constitutes an adult movie? Create a table of all movies rated TV-MA or R

adultmovies <- netmovies %>% filter(rating=="TV-MA" | rating=="R")

# Most common genres with an R or TV-MA rating
adultmovies %>% separate_rows(genres,sep=",") %>% group_by(genres) %>% summarize(movies=n()) %>% filter(genres != "" | !is.na(genres)) %>% arrange(desc(movies)) %>% head(n=10)

# Most common keywords in movies with an R or TV-MA rating
adultmovies %>% separate_rows(keywords,sep=",") %>% group_by(keywords) %>% summarize(movies=n()) %>% filter(keywords != "" | !is.na(keywords)) %>% arrange(desc(movies)) %>% head(n=10)

# Most common words found in descriptions of R or TV-MA rated movies
moviedescriptions <- adultmovies %>% separate_rows(description, sep=" ") %>% group_by(description) %>% summarize(movies=n()) %>% filter(description != "" | !is.na(description)) %>% arrange(desc(movies)) %>% head(n=100)

```

Identifying the characteristics of an adult-rated show:

```{r}
# What constitutes an adult show? Create a table of all shows rated TV-MA or R
adultshows <- netshows %>% filter(rating=="TV-MA" | rating=="R")

# Most common genres in shows with an R or TV-MA rating
adultshows %>% separate_rows(genres,sep=",") %>% group_by(genres) %>% summarize(shows=n()) %>% filter(genres != "" | !is.na(genres)) %>% arrange(desc(shows)) %>% head(n=10)

# Most common words found in descriptions of R or TV-MA rated shows
showdescriptions <- adultshows %>% separate_rows(description, sep=" ") %>% group_by(description) %>% summarize(shows=n()) %>% filter(description != "" | !is.na(description)) %>% arrange(desc(shows)) %>% head(n=100)
```

Looking at a distribution of the ratings over time, we can see below that Netflix chose to *expand on content appropriate for adults and teenagers (not suitable for ages under 14).* 

```{r}
# Plotting the ratings of Netflix movie releases over time 

netmovies %>% group_by(rating, year_added) %>% summarize(movies=n()) %>%  filter(rating != "" & year_added > 2015 & year_added < 2022 & movies > 2) %>% ggplot(aes(x=year_added,y=movies,fill=rating)) + geom_area(stat="identity") + xlim(2016,2021) + labs(title="Netflix Movie Ratings from 2016-2021")

```

```{r}
# Plotting the ratings of Netflix TV show releases over time
netshows %>% group_by(rating, year_added) %>% summarize(shows=n()) %>%  filter(rating != "" & year_added > 2015 & year_added < 2022 & shows > 2) %>% ggplot(aes(x=year_added,y=shows,fill=rating)) + geom_area(stat="identity") + xlim(2016,2021) + labs(title="Netflix TV Show Ratings from 2016-2021")
```

### Analyzing Keywords of Netflix Content

Keywords is a column that identifies common words or phrases associated with a specific movie or TV show. Studying the most frequently occurring keywords, or even its correlations with metrics such as popularity or audience scores can be an approximate measure of user preference (and thus performance)

```{r}
# Creating a table of the most commonly used keywords and their average audience scores for Netflix movies

moviekeywords <- netmovies %>% separate_rows(keywords, sep=",") %>% group_by(keywords) %>% summarize(total_votes=sum(vote_count),mean_score=mean(vote_average), avg_revenue=mean(revenue),avg_release_yr=mean(release_year),freq=n()) %>% arrange(desc(freq)) %>% filter(keywords != "" & !is.na(keywords))

head(moviekeywords)
```

The most common keywords amongst Netflix titles are **stand-up comedy, woman director, based on novel or book, murder, or based on true story.**  

With the plots below, it's evident that the commonly preferred movies on Netflix are *more popular (high engagement from audiences/high vote counts) than good (high audience scores).*  

```{r}
# Plotting the correlation of common keywords with the movie's average audience score
ggplot(data=moviekeywords) + geom_point(mapping=aes(x=mean_score,y=freq))

# Plotting the correlation of common keywords with a movie's popularity
ggplot(data=moviekeywords) + geom_point(mapping=aes(x=freq,y=total_votes))
```
